---
title: "p8105_hw6_yc4018"
author: "Yuxuan Chen"
date: "12/3/2021"
output: github_document
---
```{r message = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis" 
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

### **Problem 1**

understand the effects of several variables on a child’s birthweight.

1. Load and clean the data for regression analysis 

```{r warning = FALSE}
birthweight_df = 
  read.csv("./data/birthweight.csv") %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace))
knitr::kable(birthweight_df[0:10,])

# check for missing data
sum(is.na(birthweight_df))
```

 - The dimension of this `birthweight_df` dataset is (`r dim(birthweight_df)`). Through data cleaning, I changed the `babysex`, `frace`, `malform`, and `mrace` these categorical variables from numeric to factor. Also, after checking for missing data, there are no missing values in this dataset.   

2. Propose a regression model for birthweight

  - After searching online, I found out that the premature birth, mother's race, age, and health, multiple birth, and whether or not she smokes are the majority causes of low birthweight. Hence, I choose to study whether birthweight as an outcome depends on gestational age in weeks (`gaweeks`), mother’s race (`mrace`), mother’s age at delivery (`momage`), mother’s pre-pregnancy BMI (`ppbmi`), number of live births prior to this pregnancy (`parity`), and average number of cigarettes smoked per day during pregnancy(`smoken`).
  
```{r}
model_fit1 = lm(bwt ~ gaweeks + mrace + momage + ppbmi + parity + smoken, data = birthweight_df)
broom::tidy(model_fit1)
```

3. plot of model residuals against fitted values

```{r message = FALSE}
birthweight_df %>% 
  modelr::add_predictions(model_fit1) %>% 
  modelr::add_residuals(model_fit1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Plot of Model1: Residuals Against Fitted Values",
    x = "Fitted values",
    y = "Residuals")

```

4. Fit two other models: 

 - One using length at birth and gestational age as predictors (main effects only)
```{r}
model_fit2 = lm(bwt ~ blength + gaweeks, data = birthweight_df)
```

 - One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
model_fit3 = lm(bwt ~ bhead + blength + babysex + bhead * blength + 
                  bhead * babysex + blength * babysex + bhead * blength * babysex, data = birthweight_df)
```

5. Compute rmse: cross validation.

```{r warning = FALSE}
cv_df = 
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    model_fit1 = map(.x = train, ~lm(bwt ~ gaweeks + mrace + momage + ppbmi + parity + smoken, data = .x)),
    model_fit2 = map(.x = train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model_fit3 = map(.x = train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + 
                                       bhead * babysex + blength * babysex + bhead * blength * babysex, data = .x))) %>% 
  mutate(
    rmse_model_1 = map2_dbl(.x = model_fit1, .y = test, ~rmse(model = .x, data = .y)), 
    rmse_model_2 = map2_dbl(.x = model_fit2, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model_3 = map2_dbl(.x = model_fit3, .y = test, ~rmse(model = .x, data = .y)))
```

6. violin plot of RMSEs 

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin() +
  labs(
     title = "Violin Plot of RMSEs for Three Models",
     x = "Model",
     y = "RMSE")
  
```

 - Based on the above violin plot of RMSEs for three models, since the model 3 has the smallest RMSE value, it is the best model; and the model 1 which with the largest RMSE value is the worst model. Hence, the model that has highest prediction accuracy is the one to use `head circumference`, `length`, `sex`, and all interactions as predictors to predict child’s birthweight. 
