---
title: "p8105_hw6_yc4018"
author: "Yuxuan Chen"
date: "12/3/2021"
output: github_document
---
```{r message = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis" 
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

**Problem 1**

understand the effects of several variables on a child’s birthweight.

1. Load and clean the data for regression analysis 

```{r warning = FALSE}
birthweight_df = 
  read.csv("./data/birthweight.csv") %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace))
knitr::kable(birthweight_df[0:10,])

# check for missing data
sum(is.na(birthweight_df))
```

 - Through data cleaning, I changed the `babysex`, `frace`, `malform`, and `mrace` these categorical variables from numeric to factor. Also, after checking for missing data, there are no missing values in this dataset.

2. Propose a regression model for birthweight

  - After searching online, I found out that the premature birth, mother's race, age, and health, multiple birth, and whether or not she smokes are the majority causes of low birthweight. Hence, I choose to study whether birthweight as an outcome depends on gestational age in weeks (`gaweeks`), mother’s race (`mrace`), mother’s age at delivery (`momage`), mother’s pre-pregnancy BMI (`ppbmi`), number of live births prior to this pregnancy (`parity`), and average number of cigarettes smoked per day during pregnancy(`smoken`).
  
```{r}
model_fit1 = lm(bwt ~ gaweeks + mrace + momage + ppbmi + parity + smoken, data = birthweight_df)
broom::tidy(model_fit1)
```

3. plot of model residuals against fitted values

```{r}
birthweight_df %>% 
  modelr::add_predictions(model_fit1) %>% 
  modelr::add_residuals(model_fit1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(
    title = "Plot of Model1: Residuals Against Fitted Values",
    x = "Fitted values",
    y = "Residuals"
    )

```




